Reducing Linearizability of Priority Queues and More Data Structures to State Reachability

### 2017-04-20 co-regular-EPQ.tex
first part of co-regular
modify

Also, we consider $\Gamma$-complete automata for $\Gamma\in \{\mathsf{UnmatchedMaxPriority}, \mathsf{MatchedMaxPriority}\}$, recognizing executions where any two occurring priorities are comparable (w.r.t. $\prec$). This is possible because any data-differentiated execution which is not $\Gamma$-linearizable has such a projection. Formally, given a data-differentiated execution $e$ and $p$ a maximal priority in $e$, $e\vert_{\preceq p}$ is the projection of $e$ to the set of values with priorities comparable to $p$. Then,

@@

A differentiated execution $e$ is not $\mathsf{MatchedMaxPriority}$-linearizable when all the $\textit{put}$ operations in $e$ using the maximal priority $p$ are matched, and $e$ is not linearizable w.r.t. the set of sequential executions satisfying $\mathsf{MatchedMaxPriority\text{-}Seq}(e,x)$ for some value $x$ of priority $p$. We consider two cases depending on whether $e$ contains exactly one value with priority $p$ or at least two values. We denote by $\mathsf{MatchedMaxPriority}^{>}$ the strengthening of $\mathsf{MatchedMaxPriority}$ with the condition that all the values other than $x$ have a priority strictly smaller than $p$ (corresponding to the first case), and by $\mathsf{MatchedMaxPriority}^{=}$ the strengthening of the same formula with the negation of this condition (corresponding to the second case). We use particular instances of register automata~\cite{finite-memory, the paper I found} whose states include only two registers, one for storing a priority guessed at the initial state, and one for storing the priority of the current action in the execution. The transitions can check equality or the order relation $\prec$ between the values stored in the two registers. Instead of formalizing the full class of register automata, we consider a simpler class which suffices our needs. More precisely, we consider a class of labeled transition systems whose states consist of a finite control part and a register $r$ interpreted to elements of $\mathbb{P}$. The transition labels can be one of the following: 
\begin{itemize}
	\item $r=*$ for storing an arbitrary value to $r$, 
	\item $\textit{call}(\textit{rm},a)$ and $\textit{ret}(\textit{rm},a)$ for reading call/return actions of a remove, 
	\item $\textit{call}(\textit{put},d,g)$ where $g\in\{=r,\prec r,true\}$ is a guard, for reading a call action $\textit{call}(\textit{put},d,p)$ of a put and checking whether $p$ is either equal to or smaller than the value stored in $r$, or arbitrary,
	\item $\textit{ret}(\textit{put},d,true)$ for reading a return action $\textit{ret}(\textit{put},d,p)$ for some $p$.
\end{itemize}
The set of words accepted by such a transition system can be defined as usual.

@@

TODO INTRODUCE THE LEFT-RIGHT CONSTRAINT

>>>>

{\color {blue}Also, we consider $\Gamma$-complete automata for $\Gamma\in \{\mathsf{UnmatchedMaxPriority}, \mathsf{MatchedMaxPriority}\}$, recognizing executions which contains only one maximal priority. This is possible because any data-differentiated execution for which $\Gamma(e)$ is false has such a projection.}
%Also, we consider $\Gamma$-complete automata for $\Gamma\in \{\mathsf{UnmatchedMaxPriority}, \mathsf{MatchedMaxPriority}\}$, recognizing executions where any two occurring priorities are comparable (w.r.t. $\prec$). This is possible because any data-differentiated execution which is not $\Gamma$-linearizable has such a projection. 
Formally, given a data-differentiated execution $e$ and $p$ a maximal priority in $e$, $e\vert_{\preceq p}$ is the projection of $e$ to the set of values with priorities comparable to $p$. Then,

@@

{\color {blue}When $\mathsf{MatchedMaxPriority}$ is false for some execution $e$, then there are two possibilities:

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] There exists value $x$, let $e_x$ be obtained from $e$ when all the values of maximal priorities in $e$ but $x$ are removed. Then $\mathsf{MatchedMaxPriority\text{-}Conc}(e_x,x)$ is false.

\item[-] There exists set $X$ of values, let $e'$ be obtained from $e$ when all the values of maximal priorities in $e$ but values in $X$ are removed. Then, although $\mathsf{MatchedMaxPriority\text{-}Conc}(e'_x,x)$ is true for each $x \in X$, there is no value $y$ to make $\mathsf{MatchedMaxPriority\text{-}Conc}(e',y)$ to be true.
\end{itemize}

It is surprised to see that these two possibilities can be captured by two sub-cases of $\mathsf{MatchedMaxPriority}$, depending on whether is only one value $x$ with maximal priority $p$. We denote by $\mathsf{MatchedMaxPriority}^{>}$ the strengthening of $\mathsf{MatchedMaxPriority}$ with this condition (i.e., all the values other than $x$ have a priority strictly smaller than $p$) and $\mathsf{MatchedMaxPriority}^{=}$ for the strengthening of the same formula with the negation of this condition (i.e., there exists another value $y$ with priority $p$).
}

%A differentiated execution $e$ is not $\mathsf{MatchedMaxPriority}$-linearizable when all the $\textit{put}$ operations in $e$ using the maximal priority $p$ are matched, and $e$ is not linearizable w.r.t. the set of sequential executions satisfying $\mathsf{MatchedMaxPriority\text{-}Seq}(e,x)$ for some value $x$ of priority $p$. We consider two cases depending on whether $e$ contains exactly one value with priority $p$ or at least two values. We denote by $\mathsf{MatchedMaxPriority}^{>}$ the strengthening of $\mathsf{MatchedMaxPriority}$ with the condition that all the values other than $x$ have a priority strictly smaller than $p$ (corresponding to the first case), and by $\mathsf{MatchedMaxPriority}^{=}$ the strengthening of the same formula with the negation of this condition (corresponding to the second case). 
We use particular instances of register automata~\cite{finite-memory, the paper I found} whose states include only two registers, one for storing a priority guessed at the initial state, and one for storing the priority of the current action in the execution. The transitions can check equality or the order relation $\prec$ between the values stored in the two registers. Instead of formalizing the full class of register automata, we consider a simpler class which suffices our needs. More precisely, we consider a class of labeled transition systems whose states consist of a finite control part and a register $r$ interpreted to elements of $\mathbb{P}$. The transition labels can be one of the following:
\begin{itemize}
	\item $r=*$ for storing an arbitrary value to $r$,
	\item $\textit{call}(\textit{rm},a)$ and $\textit{ret}(\textit{rm},a)$ for reading call/return actions of a remove,
	\item $\textit{call}(\textit{put},d,g)$ where $g\in\{=r,\prec r,true\}$ is a guard, for reading a call action $\textit{call}(\textit{put},d,p)$ of a put and checking whether $p$ is either equal to or smaller than the value stored in $r$, or arbitrary,
	\item $\textit{ret}(\textit{put},d,true)$ for reading a return action $\textit{ret}(\textit{put},d,p)$ for some $p$.
\end{itemize}
The set of words accepted by such a transition system can be defined as usual.

@@

TODO GIVE THE FORMAL DEFINITION OF THE LEFT-RIGHT CONSTRAINT
